---
title: "Access PatientProfiler CPTAC Data"
author: "PerfettoLab"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

PatientProfiler is an R package that allows personalized analysis of an
entire cohort and specific individuals. The main functions can be used
sequentially or separately.

If we want to imagine a sequential way to use PatientProfiler, we can
divide the pipeline into 5 main steps:

-   **STEP 1** - Raw data harmonization/Access to harmonized CPTAC
    data & Input tables preparation

-   **STEP 2** - Protein activity inference

-   **STEP 3** - Generation of mechanistic models

-   **STEP 4** - Network-based stratification (communities detection)

-   **STEP 5** - Identification of biomarkers

![](images/Figure1-2.svg)

In this vignette, we describe how to access pre-harmonized CPTAC data to use them in PatientProfiler pipeline.

## [Load necessary libraries]{style="color: #A45C40;"}

```{r eval=FALSE}
library(PatientProfiler)
library(SignalingProfiler)
library(tidyverse)
```

## [ACCESS TO PRE-HARMONIZED CPTAC DATA]{style="color: #41729F;"}

### *access_harmonized_CPTAC_data* {style="color: black"}

If you want to use *pre-harmonized data* for running PatientProfiler, you can access them 
using the following function. You can specify one or a set of tumor types. 
start with this function of PatientProfiler.
Our processed database includes transcritpomics, proteomics and phosphoproteomics data
for 10 different tumors:

-   **Brca** - Breast cancer

-   **Ccrcc** - Clear cell renal cell carcinoma

-   **Coad** - Colon adenocarcinoma

-   **Gbm** - Glioblastoma

-   **Hnscc** -Head and neck squamous cell cacinoma

-   **Lscc** - Lung squamous cell carcinoma

-   **Luad** - Lung cancer

-   **Ov** - Ovarian cancer

-   **Pdac** - Pancreatic ductal adenocarcinoma

-   **Ucec** - Uterine corpus endometrial carcinoma

All the data has been harmonized using [omics_update] function.
The function returns an RDS file containing CPTAC pre-harmonized data for user-selected tumor types.

### [Example of usage:]{style="color: #C38370;"}

If you need to extract **Brca** and **Ov** **phosphoproteomics** and
**proteomics** data:

```{r chunk3, eval = FALSE}
access_harmonized_CPTAC_data(tumors = c("Brca", "Ov"), omics = c("phospho", "prot"))
```

The output is a local variable for each data frame you acceded.

## [NEXT STEPS IN PatientProfiler]{style="color: #41729F;"}
You can follow the main vignette to use the CPTAC pre-harmonized data and generate patient-specific networks.
